\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-04-21 sections/results.tex
Fix: Section related to Figure 4

  To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
  In both the comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
  The values shown in the figures correspond to the mean UniFrac values across all samples in the study.
  The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
  From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences of low abundance.

  % FIXME: Supplementary work - Figure S3
  In order to verify this claim, for each of these methods we use only the top 100 reference sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
  We observe that the both the weighted and unweighted UniFrac distances are decreased considerably implying that the top reference sequences generated by the different methods are similar to each other.

  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets.
  In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
  We next compared the predicted representative sequences with the expected representative sequences and their distribution.
  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
  This result is more exaggerated in the case of the unweighted UniFrac metric, where some of the datasets show a very high deviation from the expected sequences.
  These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
  Open-reference and denovo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
  This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
  Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
  The Deblur method returns a very small number of reference sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
  There is no method that clearly outperforms the rest in all datasets.
  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur seem to be in general the most reliable.
  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.

@@

  The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
  Figure~\ref{fig:figure4}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
  We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
  % TODO: Annotate Figure 4A with bacteria names
  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
  Even the most abundant representative sequence is assigned to an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
  We also observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences.
  For example, the most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
  Although these are minor differences, when comparing a large number of taxonomy composition tables these problems are hard to diagnose.

  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<67\%$ similarity between any pair of databases.
  % TODO: Verify this
  The assignment between SILVA and the other two databases is significantly different ($40\%$ mismatch) even at the Phylum level, indicating that 40 of the top 100 assignments at the Phylum level in the SILVA database are different in SILVA.
  The \ac{gg} and \ac{ncbi} assignments are more similar compared to the other pairwise comparisons, mainly due to the poor performance of SIVLA at the Phylum level.
  However, the number of mismatches for taxonomy assignments beyond the Family level are very high ($>66\%$), implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.

>>>>

  To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
  In both the comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
  The values shown in the figures correspond to the mean UniFrac values across all samples in the study.
  The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
  From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences of lower abundance.

  In order to verify this claim, for each of these methods we use only the top 1000 representative sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
  We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not similar to each other.
  % TODO: Can we make this conclusion?
  This leads us to the conclude that the representative sequences in the middle range of the abundance distribution are those that are most similar between the methods.

  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets.
  In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
  We next compared the predicted representative sequences with the expected representative sequences and their distribution.
  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
  This result is more exaggerated in the case of the unweighted UniFrac metric, where some of the datasets show a very high deviation from the expected sequences.
  These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
  Open-reference and denovo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
  This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
  Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
  The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
  There is no method that clearly outperforms the rest in all datasets.
  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur seem to be in general the most reliable.
  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.

@@

  The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
  We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
  The large percentage of unassigned sequences indicate that even among the most abundant representative sequences a large fraction is assigned an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proporitions in the alluvial plot).
  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{gg} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in NCBI and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
  Although these are minor differences (in the case of \ac{gg} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.

  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<45\%$ similarity between any pair of databases.
  The matches and mismatches in the assignments from all three databases are comparable, with \ac{ncbi} and SILVA being the most similar across most of the taxonomic levels, followed closed by \ac{gg} vs. SILVA.
  However, the number of mismatches for taxonomy assignments beyond the Genus level (Genus, Species) are very high ($>67\%$), implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
  The assignment between SILVA and the other two databases were originally significantly different ($40\%$ mismatch) even at the Phylum level, indicating that 40 of the top 100 assignments at the Phylum level in the SILVA database are different in SILVA.
  However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.

### 2022-04-21 sections/supplementary.tex
Add: Preliminary scripts for Figure S4
Fix: Section related to Figure 4

>>>>

    % TODO: Write something about TSNE
    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The TSNE plot of the networks inferred by various methods}.
        }
      \label{fig:figure_s2}
    \end{figure}
    \FloatBarrier
    \newpage

  \subsection*{Similarity of high abundance reference sequences}

    Figure~\ref{fig:figureS3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
    In Figure~\ref{fig:figureS3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
    We observe that both the weighted and unweighted UniFrac distances are increased compared to the figure with the all the representative sequences.
    This shows that the most abundant representative sequences generated by the methods are not similar to each other.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The UniFrac distance between the 100 most abundant representative sequences generated by various method in the \ac{dc} step}
        }
      \label{fig:figure_s3}
    \end{figure}
    \FloatBarrier
    \newpage

    % TODO: Figure S4 - Unifrac bar plot for CC methods
    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
        }
      \label{fig:figure_s4}
    \end{figure}
    \FloatBarrier
    \newpage

  \subsection*{}
